\tw: Building a Xanalogical System in Ten Days

### 2003-02-13 papers/hypertext2003/paper.tex
Added a related work section, and another reference.

Much of the existing work on xanalogical systems was discussed in Sections \ref{sec:Introduction} and \ref{sec:Xanadu}.

Wikis
\cite{WikiWikiWeb}
\cite{Wikipedia}

Everything2
\cite{Everything2}

Weblogs
\cite{kuro5hin}

Xanadu

>>>>

Much of the existing work on xanalogical systems was discussed in Sections \ref{sec:Introduction} and \ref{sec:Xanadu}, so we will focus on related non-xanalogical systems here.
However, we should mention that interest in xanalogical systems has not waned in recent years.
For example, Lukka and Fallenstein describe a system that uses ideas taken from Freenet for xanalogical content storage and retrieval \cite{FreenetXanadu}.

\subsection{Wikis}

In terms of implementation, \tw \ is closely related to a recent class of hypertext systems known as Wikis, which are often implemented as Perl-based CGI programs.
Wikis are reader-edited web sites, and web-based HTML forms are used as the interface for page authoring.
In many cases, Wikis are open for editing by the public at large.

Aside from the sociological issues, Wikis approach hypertext in an interesting way.
Links can be created in a document to documents that do not exist without knowing where those documents will eventually reside.
To create a new document, a user must follow one of these dangling links in an existing page, which leads to a page creation screen.
Thus, in a Wiki, there is no way to create a page in isolation.
In fact, the name of a new page is determined by the contents of the dangling link and cannot be specified in the page creation screen.

\tw \ differs from a Wiki in its approach to hypertext, since a \tw \ quote can only be taken from an existing document.
No mechanism is provided for refering to a document that does not yet exist.

The first Wiki was created by Ward Cunningham in 1995 and is still active today\cite{WikiWikiWeb}.
The most ambitious Wiki-based project is Wikipedia, a user-edited encyclopedia, which recently grew to over 100,000 entries \cite{Wikipedia}.

\subsection{Everything2}
Like \tw \ and Wikis, Everything2 is a web-based hypertext system implemented using CGI and Perl \cite{Everything2}.
Everything2 supports dangling links to pages that do not exist yet, but it handles these links differently than a Wiki does.
When a user follows a dangling link, a search is performed for the title of the non-existant page.
Thus, dangling links are fuctional from a reader's point of view.
We might conclude that Wikis favor authors, while Everything2 favors readers.
As soon as a page is created that matches the title of a dangling link, that link points to the new page directly.

In addition to in-line, user-created links (dangling or otherwise), each node in Everything2 features a list of automatically generated links.
These links are culled from user traversal paths:  when a user leaves one page and enters another (either by clicking a link or performing a search), a two-way link is added between the pages.

\subsection{Weblogs}
Another class of related web based hypertext systems are weblogs, or ``blogs.''
These systems allow content to be added to a web site serially, one post at a time.
Some blogs are used by individuals as a means of sharing daily thoughts with the world, while others aim to be collaborative information spaces filled with moderated comments posted by the public.
One of the most interesting collaborative weblogs is Kuro5hin (pronounced ``corrosion''), which features high-quality news and culture stories submitted and moderated by the public \cite{kuro5hin}.
Many blogs are implemented using CGI and Perl, though more recent efforts seem to favor PHP.

\subsection{Commonalities and Differences}
All three systems share several aspects in common with \tw.
First, they are all web-based, collaborative information spaces.
Second, they all relie heavily on existing technology for their implementations.
Third, from what we can tell, they were developed relatively quickly.

In addition, we can see iterative development philosophies in weblogs and Wikis, not necessarily in the individual projects, but in the phenomena themselves.
The first Wikis and weblogs were simple but functional.
Over time, new implementations were released with more features.
Today, the most popular systems in either category are extraordinarily feature-rich.

Of course, out of all of the collaborative information spaces discussed here, \tw \ is the only system that supports deep content reuse and frictionless micropayments.
Thus, these systems are closely related in terms of implementation, but their goals are very different.
